---
title: "p8105_hw1_zx2528"
author: "Zizhi Xia"
date: "2025-09-20"
output: github_document
editor_options:
  chunk_output_type: inline
---

# Problem 1

First, we load the dataset:
```{r}
library(tidyverse)
library(moderndive)
data("early_january_weather")
```
The dataset `early_january_weather` has `r nrow(early_january_weather)` rows
and `r ncol(early_january_weather)` columns. 
The mean temperature is `r round(mean(early_january_weather$temp, na.rm = TRUE), 1)` degrees Fahrenheit.  
Important variables include `r paste(names(early_january_weather), collapse = ", ")`.

```{r}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point(alpha = 0.6) +
  labs(
    title = "Temperature vs Time in Early January",
    x = "Time",
    y = "Temperature (F)",
    color = "Humidity"
  )
```
```{r}
ggsave("scatterplot_temp_time.png", width = 8, height = 6)
```
From the scatterplot, we see that temperature generally fluctuates over time in early January, with a noticeable upward trend around mid-January. Humidity, shown by the color scale, is typically higher (darker blue) when temperatures are lower, and decreases as temperatures rise.

# Problem 2

```{r}
set.seed(8105)  
n <- 10

df <- tibble::tibble(
  num   = rnorm(n),                                  
  logic = num > 0,                                    
  char  = sample(letters[1:5], n, replace = TRUE),    
  fac   = factor(sample(c("low","mid","high"), n, TRUE),
                 levels = c("low","mid","high"))      
)

df
```
```{r}
library(tidyverse)

mean_num   <- df %>% pull(num)   %>% mean()
mean_logic <- df %>% pull(logic) %>% mean()
mean_char  <- try(df %>% pull(char) %>% mean(),  silent = TRUE)
mean_fac   <- try(df %>% pull(fac)  %>% mean(),  silent = TRUE)

tibble(
  variable   = c("num","logic","char","fac"),
  mean_value = c(mean_num, mean_logic, NA, NA),
  note = c(
    "OK: numeric",
    "OK: logical ",
    "ERROR: character ",
    "ERROR: factor "
  )
)
```

```{r}
as_num_logic <- as.numeric(df$logic)  
as_num_char  <- as.numeric(df$char)   
as_num_fac   <- as.numeric(df$fac)  
```
This exercise shows that numeric and logical variables can be meaningfully averaged, 
while character and factor variables cannot. 
